### 2025-04-05 decision-making.qmd
20250405 - additions

As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case, `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy ( MdAE = `{r} mdaeBracketingnotCentered2_individual`).

@@

## Strategies for Managing Risk and Uncertainty {#sec-managinRiskUncertainty}

@@

Fantasy football (and sports betting, more generally) is not 100% luck.
Skill can be involved in identifying undervalued assets—whether in terms of stocks in the stock market or professional football players, where inside information can be especially valuable (and thus illegal to profit from).
To evaluate the percentage of variability in fantasy football that is attributable to luck versus skill, @Getty2018 evaluated the extent to which one's performance in the first half of the season was correlated with one's performance in the second half of the season, under the assumption that underlying skill would lead to persistence of performance across time.
The authors examined FanDuel daily fantasy sports (DFS) leagues, which allow the user to set an entirely new lineup each week from all players, based on salary constraints ( so a user cannot just select the best player at every position).
The authors estimated that ~55% of the variability in fantasy football performance across time was due to skill and 45% was due to luck [@Getty2018].
Thus, performance in fantasy football is around half and half luck versus skill.

@@

- seek advice from diverse perspectives and incorporate it into your decision making
- try to get opinions from others *before* you state your perspective and before the various sources of advice discuss with each other, to ensure independence of advice
  - As noted by @Kahneman2011, "before an issue is discussed, all members of the committee should be asked to write a very brief summary of their position... The standard practice of open discussion gives too much weight to the opinions of those who speak early and assertively, causing others to line up behind them." (p. 85).
- if some sources of advice (or projections) are clearly more skilled and accurate than others, you can average this "select" crowd of projections or give them greater weight in a weighted average
- if it is unclear whether or which sources are reliably more accurate than others, using a simple average across all sources (i.e., crowd-averaged projections) can be a useful approach that is as accurate as—if not more accurate than—the average individual forecaster
- incorporate at least 5–10 sources of projections

>>>>

As demonstrated above, when the individual predictions bracket but are not centered on the truth, and the bracketing rate is high (in this case, `{r} petersenlab::apa(bracketingnotCentered2_bracketingRate, decimals = 2)`%) but with an outlier projection that is far away from the truth, the crowd-averaged predictions (MAE/MdAE = `{r} maeBracketingnotCentered2_crowd`) are more accurate than the accuracy of the typical individual forecaster (MAE = `{r} maeBracketingnotCentered2_individual`) in terms of [mean absolute error](#sec-meanAbsoluteError), but they can still be less accurate—in terms of [median absolute error](#sec-medianAbsoluteError)—than the individual forecaster who is at the 50th percentile in accuracy (MdAE = `{r} mdaeBracketingnotCentered2_individual`).

@@

## Strategies for Managing Risk and Uncertainty {#sec-managingRiskUncertainty}

@@

The question of whether fantasy football is a game of chance or skill is important because such considerations help determine whether betting on one's performance is considered gambling, which might be illegal or regulated in many jurisdictions.
Fantasy football (and sports betting, more generally) is not 100% luck.
Skill can be involved in identifying undervalued assets—whether in terms of stocks in the stock market or professional football players, where inside information can be especially valuable (and thus illegal to profit from).
To evaluate the percentage of variability in fantasy football that is attributable to luck versus skill, @Getty2018 evaluated the extent to which one's performance in the first half of the season was correlated with one's performance in the second half of the season, under the assumption that underlying skill would lead to persistence of performance across time.
The authors examined FanDuel daily fantasy sports (DFS) leagues, which allow the user to set an entirely new lineup each week from all players, based on salary constraints (so a user cannot just select the best player at every position).
The authors estimated that ~55% of the variability in fantasy football performance across time was due to skill and 45% was due to luck [@Getty2018].
Thus, performance in fantasy football is around half and half luck versus skill.

@@

- Seek advice from diverse perspectives and incorporate it into your decision making.
- Try to get opinions from others *before* you state your perspective and before the various sources of advice discuss with each other, to ensure independence of advice.
  - As noted by @Kahneman2011, "before an issue is discussed, all members of the committee should be asked to write a very brief summary of their position... The standard practice of open discussion gives too much weight to the opinions of those who speak early and assertively, causing others to line up behind them." (p. 85).
- If some sources of advice (or projections) are clearly more skilled and accurate than others, you can average this "select" crowd of projections or give them greater weight in a [weighted average]().
- If it is unclear whether or which sources are reliably more accurate than others, using a simple average across all sources (i.e., crowd-averaged projections) can be a useful approach that is as accurate as—if not more accurate than—the average individual forecaster.
- Incorporate at least 5–10 sources of projections.
Use a [weighted](#sec-descriptiveStatisticsCenter) or [robust](#sec-descriptiveStatisticsCenter) average to account for outlier projections.
- Do not bet on fantasy football (or anything else for that matter) unless you are willing to lose the money.
Sports bettors tend to be [overconfident](#sec-cognitiveBiasesOverconfidence); on average, they lose 7.5 cents for every dollar wagered [@Brown2025].
If you are going to bet, only bet a small portion of your money and never more than you are willing to lose.
And do not make multi-leg bets (parlays); people lose on average 25 cents for every dollar wagered on parlays [@Brown2025].
